target_offtarget

### 2025-01-30 manuscript_figures.Rmd
Figures S10-S12 from genom med paper

ggsave("figures/manuscript_figure_2025/PNG/Figure_3.png", 
       width = 10, height = 8)
```

### ----nodedup results ----
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=6}
##extras - no dedup

nodedup_reads <-
  nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(count),
    colour = genus
  )) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  ylab("log10(Viral Reads)") +
  scale_color_manual(values = cols)

nodedup_prop_viral <- 
  nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of viral reads") +
  scale_color_manual(values = cols)

nodedup_prop_all <- nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_total, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of total reads") +
  scale_color_manual(values = cols)

ggarrange(
  nodedup_reads,
  nodedup_prop_all,
  nodedup_prop_viral,
  nrow = 3,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/target_offtarget_nodedup.pdf",width=8,height=10)

@@

>>>>

ggsave("figures/manuscript_figure_2025/PNG/Figure_3.png", 
       width = 10, height = 8)
```
### --------------------

### Figure S10

#### Read-in data
```{r}

gm_readcount_dedup <-
  read.table(
    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

gm_readcount_nodedup <-
  read.table(
    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

gm_counts_all <- rbind(gm_readcount_dedup, gm_readcount_nodedup)

@@

#### read counts 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#normalise by both raw read count and genome length - should be the same as mean read depth

facet_names <- c("dedup_GM_TE" = "Deduplicated",
                 "nodedup_GM_TE" = "Non-Deduplicated")

gm_counts_norm <- 
  gm_counts_all |>
  mutate(norm_counts3 = matched / length)

read_counts <- 
  gm_counts_norm |>
  filter(Viral.load != 0) |>
  ggplot(aes(
    x = Viral.load,
    y = (matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus), se = FALSE, span = 0.85) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    # axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Viral Reads") +
  xlab("Viral load (gc/ml)") + 
  scale_x_log10(labels = scales::trans_format("log10",
                                              scales::label_math())) + 
  scale_y_log10(labels = scales::trans_format("log10",
                                              scales::label_math()))

#### read depth
```{r}
#import and combine read depth files

depth_dedup_gm <-
  read.table(
    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depth_nodedup_gm <-
  read.table(
    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depths_gm_all <- rbind(depth_dedup_gm, depth_nodedup_gm)

```{r, echo=FALSE, warning=FALSE, message=FALSE}

read_depths <- 
  depths_gm_all |>
  filter(Viral.load != 0) |>
  ggplot(aes(
    x = Viral.load,
    y = mean_depth,
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus), se = FALSE, span = 0.9) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    # axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Mean read depth") +
  xlab("Viral load (gc/ml)") + 
  scale_x_log10(labels = scales::trans_format("log10",
                                              scales::label_math())) + 
  scale_y_log10(labels = scales::trans_format("log10",
                                              scales::label_math()))

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggarrange(read_counts,
          read_depths,
          nrow = 2,
          common.legend = TRUE)

#PDF
ggsave("figures/manuscript_figure_2025/PDF/Figure_S10.pdf", 
       width = 7,
       height = 7)

#PNG
ggsave("figures/manuscript_figure_2025/PNG/Figure_S10.png", 
       width = 7,
       height = 7)

### Figure S11

#### Read-in data
```{r}
####genome coverage####

unzip(zipfile = "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv.zip", exdir = "data_genome_medicine/")

dedup_per_site <-
  read.table(
    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

file.remove("data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv")

#View(dedup_per_site)

lengths <- 
  counts_dedup_bt |>
  select(virus,length) |>
  distinct()

coverage <- 
  dedup_per_site |>
  group_by(sample_id, virus, Viral.load, Replicate) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage / length)

coverage |>
  filter(Viral.load != 0) |>
  ggplot(aes(x = virus, y = percent_coverage)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_dodge(width = .75)) +
  facet_grid( ~ as.character(Viral.load)) +
  theme_bw() +
  theme(axis.title.y = element_text(size = 10)) +
  ylab("Genome coverage") +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 15),
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
  ) +
  scale_y_continuous(limits = c(0, 1)) +
    scale_x_discrete(
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  )

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/genome_medicine/all_genome_coverage.png",width=14,height=6)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS12.pdf",width=20,height=6)
```

#### *** Final plot ***
```{r}
#PDF
ggsave(
  "figures/manuscript_figure_2025/PDF/Figure_S11.pdf",
  width = 14,
  height = 6
)

#PNG
ggsave(
  "figures/manuscript_figure_2025/PNG/Figure_S11.png",
  width = 14,
  height = 6
)

### Figure S12
```{r, echo=FALSE, message=FALSE, warning=FALSE}
####per site plot AdV only####

dedup_per_site |>
  filter(virus == "Human_adenovirus_40") |>
  filter(Viral.load != 0) |>
  ggplot(aes(x=site,y=coverage))+
  geom_col()+
  facet_grid(Viral.load~Replicate)+
  ylab("Coverage per site")+
  ggtitle("Human Adenovirus (deduplicated)")+
  theme_few() +
  theme(
    # axis.title.x = element_text(size = 12),
    # # axis.text.x = element_text(angle = 30, hjust = 1),
    # axis.title.y = element_text(size = 12),
    # legend.title = element_blank(),
    # legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
  ) +  
  scale_y_log10(labels = scales::trans_format("log10",
                                              scales::label_math())) +

  xlab("site position")

#### *** Final plot ***
```{r}
ggsave("figures/manuscript_figure_2025/PDF/Figure_S12.pdf",
       width=12,
       height=10)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S12.png",
       width=12,
       height=10)
```
